%!TEX root = ../../main.tex
\chapter{Zero-Dose Extrapolation}
\label{chap:Zero-Dose Extrapolation}

\section{Introduction}
\label{sec:Introduction - Zero-Dose extrapolation}
Mathematical correction of reflection intensities to account for radiation damage has been a common procedure for decades \cite{hendrickson1973,abrahams1987anisotropy}.
Before the development of RADDOSE and RADDOSE-3D, the progression of radiation damage was tracked with time as the independent variable.
Functions to describe the intensity loss took the form of the Blake and Phillips model detailed in equation \ref{eq:Blake and Phillips Radiation Damage Model}, or isotropic polynomial expressions \cite{abrahams1987anisotropy}.

Radiation damage correction is still an integral part of current scaling methods \cite{otwinowski2003multiparametric,evans2005,kabsch2010integration}.
However, these correction models only compensate for the average decay of reflection intensities.
In the correction algorithms, account is not taken of specific changes.
Diederichs \textit{et al.}  introduced a linear model to perform zero-dose extrapolation, which was tested on data collected on bovine brain tubulin at 15$\,$K \cite{diederichs2003}, and later introduced exponential and quadratic models \cite{diederichs2006}, to account for these specific changes.
These authors showed that zero-dose extrapolation can improve phasing statistics and SAD electron density maps.
The problem is that reflections exhibit a diverse range of behaviours that are yet to be accounted for by a single Mathematical relationship \cite{blake1962,abrahams1973}.

Over the last decade there have been several developments in the methods and understanding of radiation damage in MX.
Some of the key advances in this field have been:
\begin{itemize}
    \item RADDOSE-3D \cite{zeldin2013}, which allows accurate 3D modelling of the diffraction experiment.
    \item the introduction of DWD \cite{zeldin2013dwd}, which is a useful metric for assessing the damage state of a crystal that resulted in the diffraction for a given image.
    \item the Leal \textit{et al.} dose decay model \cite{leal2012}. This relatively simple model was shown to accurately describe the intensity decay of reflections at RT.
\end{itemize}
The work described in this chapter is concerned with using the developments presented above to perform zero-dose extrapolation on data collected from one of the cubic crystals (crystal ID 0259) described in chapter \ref{chap:Dose Decay Modelling}.
