\newenvironment{acknowledgements}%
{\cleardoublepage\thispagestyle{empty}\null\vfill\begin{center}%
\bfseries Acknowledgements\end{center}}%
{\vfill\null}
\begin{acknowledgements}
\small
\singlespacing
First and foremost I thank my supervisor, Professor Elspeth Garman, for seeing my potential and giving me the opportunity to get to the position I am in now.
The support she has given me over the last four years is second to none, and she has only ever put me in a position to succeed.

Markus Gerstel is the biggest reason for my technical development.
Without him I would be at least six months behind with my ninja programming skills (as Oli puts it).
Markus is a great role model and I have definitely looked up to him throughout my DPhil.
The constant scientific discussions I have had with Charlie Bury have made my DPhil even more enjoyable.
He knows what questions to ask that really rack my brain.
Furthermore, he has always been supportive and contributes to my work and development in many ways.
Katharina has always been supportive and helps me with anything in the wet lab that I have little knowledge about (and that is a lot).
I will never forget how stressful it is to extract crystals from glass LCP plates.
Oli Zeldin has been a positive influence on me since day one.
He and Markus first encouraged me to join the group and I have never looked back.
To this day, Oli still believes in me and my ability and I am incredibly lucky to have him as a friend.
Much of my crystallographic knowledge I owe to Edward Lowe.
Is there anything he doesn't know?
I can ask him anything and he will always take the time out to explain it to me, as well as help me with my data processing.

I also thank the many academics that have given their support outside of the lab.
Professor Colin Please who has been a fantastic teacher and mentor ever since I began my maths degree (all those years ago).
Garib Murshudov who is the major influence on the work I've carried out over the past year.
Harry Powell, Arwen Pearson, James Holton and Andrea Thorn for great discussions and encouragement over the years.

None of this would be possible without the love and support from my family, without whom, this would not be possible.
My mother has given everything I could ever need in life and has moulded me into the man I am today.
My sister for her ongoing love, support and encouragement that has never stopped.
I could not ask anymore from my godfather who has taken me under his wing and is the perfect father figure that I will look up to for the rest of my life.
My partner, Beth Phipps, for always being by my side through the good times and the tough times.
She never stops believing in me.
I am forever indebted to them all.

Finally, I am grateful to the Systems Biology program of the EPSRC funded Doctoral Training Centre for funding the studentship that made all of this work possible.
\end{acknowledgements}
